---
Title:  Analysis for the manuscript, "RNA-Sequencing data for 7706 tumor samples across 20 cancer types from The Cancer Genome Atlas "
author: "Mumtahena Rahman"
date: "March 5, 2015"
output: html_document
---
```{r echo=FALSE}
start_time <- Sys.time()
```
## Analysis of data as described in the manuscript "RNA-Sequencing data for 7706 tumor samples across 20 cancer types from The Cancer Genome Atlas"" 

#### Preliminary steps

* Manually download the RNA-Seq data for GFP and HER2 overexpressed human mammary epithelial cells from GEO (accession# GSE62820)
* Manually download the compiled RNA-Seq and clinical data from GEO (accession# GSE62944). The Scripts/process_tcga_rsubread script contains code that shows how we processed the RNA-Seq data. The Codes/ProcessClinicalData.R script in our repository contains the code that we used to generate the clinical file. Please these in the Analysis_datasets directory.
* Install and load the required R packages for analysis: "stats", "ROCR", "pROC". The following code demonstrates how to do this.
* Set the working directory in R to the ```TCGA_RNASeq_clinical/Analysis_datasets``` directory in this repository.

```{r include=TRUE}

if (!require("stats")) {
   install.packages("stats", dependencies = TRUE)
   library(stats)
   }
if (!require("ROCR")) {
   install.packages("ROCR", dependencies = TRUE)
   library(stats)
   }
if (!require("pROC")) {
   install.packages("pROC", dependencies = TRUE)
   library(stats)
   }
```

The following code reads data files that have been preprocessed using both pipelines.

```{r "Reading counts file" ,echo=TRUE}
# TCGA pipeline, expected counts
rsem_her2_expected_counts<-read.table("GFP18_HER2_TCGA_Pipeline_Expected_Gene_Counts.txt", sep='\t', header=1, row.names=1, check.names=F) # This was downloaded from GEO Accession # GSE62820 and unzipped

# Rsubread pipeline, gene counts
feature<-read.table("GFP18_HER2_Rsubread_geneCounts.txt", sep='\t',header=1, row.names=1, check.names = F) # This was downloaded from GEO Accession # GSE62820 and unzipped

# TCGA pipeline, normalized expression files
TCGA_her2<-read.table("GFP18_HER2_TCGA_Pipeline_Normalized_Genes_Results.txt", sep='\t', header=1, check.names=F) # This was downloaded from GEO Accession # GSE62820 and unzipped

# Rsubread pipeline, FPKM values
rsub_fpkm<-read.table("GFP18_HER2_Rsubread_FPKM.txt", sep='\t',header=1, row.names=1, check.names = F) # This was downloaded from GEO Accession # GSE62820 and unzipped
rsub_fpkmlog<-log2(rsub_fpkm+1)

# Rsubread pipeline, TPM values
rsub_tpm<-read.table("GFP18_HER2_Rsubread_TPM.txt", sep='\t',header=1, row.names=1, check.names = F) # This was downloaded from GEO Accession # GSE62820 and unzipped
rsub_tpmlog<-log2(rsub_tpm+1)

# Clinical data
clinicals<-t(read.delim('TCGA20_clinical_data_ordered_all_clinical_variables_samples_as_columns.txt',sep='\t',header=1, row.names=1,check.names=F)) # This was downloaded from GEO Accession # GSE62820 and unzipped
```

We used BinReg 2 to make HER pathway predictions. The following code reads the resulting prediction files into R.

```{r}
rsub_preds<-read.table("rsubread_10_14.txt", sep='\t', header=1, row.names=1)
tcga_preds<-read.table("Rsem_10_14.txt", sep='\t', header=1, row.names=1)
```

We calculate the number of zero values per sample across the TCGA data set. The Codes/numZero.R script was used to generate the files containing the total number of zero expressing genes per samples.

```{r}
pancan12_zero<-read.table("PANCAN12_19583_by_3380_numZeroes.txt",row.names=1,sep='\t')# File is at Analysis_datasets
pancan20_tpm_zero<-read.table("PANCAN20_19583_by_3380_numZeroes.txt",sep='\t',row.names=1)# File is at Analysis_datasets
```

We used Random Forests regression to predict lung-cancer type based on RNA-Seq data. The R script for generating the predictions for this analysis is located at Code/Classify_luad_vs_lusc.R. LUAD and LUSC TCGA Level 3 data was downloaded from TCGA data portal and Rsubread TPM normalized LUAD and LUSC data was used in this analysis.

```{r}
data12 = read.table("Classification_12_LUAD_LUSC_Predictions.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE, row.names=1) # File is at Analysis_datasets
data20 = read.table("Classification_20_LUAD_LUSC_Predictions.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE, row.names=1)# File is at Analysis_datasets
```

## Helper functions

```{r}
#This function calculates the standardized mean using Hedge's formula
standardized_mean<-function(m.1,sd.1,n.1,m.2,sd.2,n.2){
  sd_pooled=sqrt(((n.1-1)*sd.1^2+(n.2-1)*sd.2^2)/(n.1+n.2-2))
  (m.1-m.2)/sd_pooled
  }

#This function merges two matrices on row names, sets the common items as rownames  and removes the extra column resulting from merge function.
merge_drop<-function(x,y,by=0)
{
  new_m<-merge(x,y,by=by)
  rownames(new_m)<-new_m$Row.names
  return(new_m[,2:length(colnames(new_m))])
}

#This function plots the ROC based on the actual and predicted class
plotROC = function(actual, probabilities, plotCI=FALSE)
{
  # bottom, left, top, right
  par(mar=c(4.5, 4.7, 0.0, 0.5),lwd=4)

  library(pROC)
  roc_result = roc(actual ~ probabilities, ci=TRUE, plot=TRUE, print.auc=FALSE)
  lowerBoundAuc = format(roc_result$ci[1], digits=3)
  midAuc = format(roc_result$ci[2], digits=3)
  upperBoundAuc = format(roc_result$ci[3], digits=3)

  if (plotCI)
  {
    ci(roc_result)
    sens.ci <- ci.se(roc_result)
    plot(sens.ci, type="shape", col="gray95")
    plot(sens.ci, type="bars")
    plot(roc_result, add=TRUE)
  }

  text(0.5, 0.00, labels=paste("AUC: ", midAuc, " (", lowerBoundAuc, "-", upperBoundAuc, ")", sep=""))

  par(mar=c(5.1, 4.1, 2.1, 2.1))
}
```

### 'Gene count' level analysis

We compared gene counts results using GFP control protein and ERBB2 protein overexpressed human mammary epithelial cell samples. TCGA's pipeline and Rsubread pipelines were used to generate raw gene level counts. Log based 2 gene counts are used to differentiate between control and ERBB2 overexpressed samples.

First, we computed the empiric cumulative distribution for all samples from the two pipelines. Second, we compared the total mapped reads per HMEC samples for the TCGA pipeline and Rsubread pipeline. Raw gene counts for GFP and HER2 HMEC samples were processed via the TCGA pipeline (expected counts) and the Rsubread pipeline (integer based), This analysis was done before any normalization method was applied.

```{r echo=TRUE}
##########computing the empiric cumulative distribution per sample overlaied on same graph######## 
###using TCGA pipelined aligned data
ecdf_all_ex<-apply(log2(rsem_her2_expected_counts+1),2,ecdf)
par( mfrow = c( 1, 2 ) )
plot(ecdf_all_ex[[1]],xlab="log2(Total mapped reads)",ylab="Cumulative proportion",col="blue",main="TCGA pipeline",ylim=c(0,1),xlim = c(0,20),cex.axis=1.5, cex.lab=1.5)
legend(10,10,c("GFP", "HER2"), col = c("blue","brown"))
for(i in 2:12){lines(ecdf_all_ex[[i]],xlab=NA, ylab = NA,col="blue")}
for(i in 13:17){lines(ecdf_all_ex[[i]],xlab=NA, ylab = NA,col="brown")}

###using Rsubread pipeline aligned data
ecdf_all<-apply(log2(feature+1),2,ecdf)
plot(ecdf_all[[1]],xlab="log2(Total mapped reads)",ylab="Cumulative proportion",col="blue",main="Rsubread pipeline",ylim=c(0,1),xlim = c(0,20),cex.axis=1.5, cex.lab=1.5)
for(i in 2:12){lines(ecdf_all[[i]],xlab=NA,ylab = NA,col="blue")}
for(i in 13:17){lines(ecdf_all[[i]],xlab=NA,ylab = NA,col="brown")}

############computing total number of read counts per samples and plotting them as dot plots####
expected_counts<-apply(rsem_her2_expected_counts,2,sum)
feature_counts<-apply(feature,2,sum)
# Creating a plot showing total mapped reads per sample  
par( mfrow = c( 1, 2 ),lwd=4 )
x = c(rep(1, 12), rep(2, 5)) # this indicates where on the x axis to plot
par(mar=c(3.1, 4.6, 2.1, 0.6)) # figure margins

boxplot(log2(expected_counts[1:12]+1), log2(expected_counts[13:17]+1),range=0,cex.axis=1.5, cex.lab=1.5,outpch=NA,lwd=4,ylim=c(20,25),xlab="", ylab="log2(Total mapped reads)",main="TCGA Pipeline",col='grey75',medcol="grey75",lwd=4,border = "grey35")

points(jitter(x, factor=2), c(log2(expected_counts[1:12]+1), log2(expected_counts[13:17]+1)), pch=4, cex=2, col=1,  xaxt="n",cex.lab=1.5)
axis(1, at=1:2, tick=T, labels=c("Control", "HER2"), cex.axis=1.5)

boxplot(log2(feature_counts[1:12]+1), log2(feature_counts[13:17]+1),range=0,cex.axis=1.5, cex.lab=1.5,outpch=NA,lwd=4,ylim=c(20,25),xlab="", ylab="log2(Total mapped reads)",col='grey75',medcol="grey75",lwd=4,main="Rsubread Pipeline",border = "grey35")
points(jitter(x, factor=1.5), c(log2(feature_counts[1:12]+1), log2(feature_counts[13:17]+1)), pch=4,cex=2,cex.lab=1.5,col="black")
axis(1, at=1:2, tick=T, labels=c("Control", "HER2"), cex.axis=1.5) 
```

Third, we are comparing the "ERBB2" gene counts across the experimental treatment to assess how well the two pipelines differentiate control ERRB2 versus overexpressed ERBB2 at the gene counts level.

```{r}
#######Boxplotting ERBB2 gene counts in HMEC samples#####
par(mfrow = c(1, 1),lwd=4)
names=c('TCGA\nGFP','TCGA\nHER2','Rsubread\nGFP', 'Rsubread\nHER2')
rsem_her2<-data.frame(t(rsem_her2_expected_counts["ERBB2",]))
rsub_her2<-data.frame(t(feature["ERBB2",]))
x = c(rep(1, 12), rep(2, 5),rep(3, 12), rep(4, 5))

boxplot(log2(rsem_her2$ERBB2[1:12]+1),log2(rsem_her2$ERBB2[13:17]+1),log2(rsub_her2$ERBB2[1:12]+1),log2(rsub_her2$ERBB2[13:17]+1),ylab="",range=0,cex.axis=1.5, cex.lab=1.5,outpch=NA,col='grey75',medcol="grey75",lwd=4,main=paste('Comparing TGCA and Rsubread Pipelines','\n', 'in Differentiating HER2 Overexpression from Controls',sep=''),border = "grey35")

points(jitter(rep(1,12),factor=2),log2(rsem_her2$ERBB2[1:12]+1),pch=4,cex=2,cex.lab=1.5,col="black")
points(jitter(rep(2,5),factor=2),log2(rsem_her2$ERBB2[13:17]+1),pch=4,cex=2,cex.lab=1.5,col='black')
points(jitter(rep(3,12),factor=2),log2(rsub_her2$ERBB2[1:12]+1),pch=4,cex=2,cex.lab=1.5,col='black')
points(jitter(rep(4,5),factor=2),log2(rsub_her2$ERBB2[13:17]+1),pch=4,cex=2,cex.lab=1.5,col='black')
axis(1, at=1:4, tick=T, labels=c("TCGA\nControl", "TCGA\nHER2","Rsubread\nControl", "Rsubread\nHER2"), cex.axis=0.8)

##using data processed by RSEM detected difference in her2 gene count in HER2 overexpressed versus GFP overexpressed samples
##t = -12.1833, df = 4.157, p-value = 0.0002081 but was worse than Rsubread
t.test(log2(rsem_her2$ERBB2[1:12]+1),log2(rsem_her2$ERBB2[13:17]+1))

##using not normalized data processed by Rsubread was much better at detecting difference in her2 gene count in HER2 overexpressed versus GFP overexpressed samples
##t = -46.6747, df = 8.35, p-value = 2.152e-11
t.test(log2(rsub_her2$ERBB2[1:12]+1),log2(rsub_her2$ERBB2[13:17]+1))

###########here we are computing standardized mean difference using the exprected gene counts from TCGA pipeline and gene counts from Rsubread algorithm ############

####Hedge's standardized mean/effect size using TCGA pipeline 
standardized_mean(mean(log2(rsem_her2$ERBB2[13:17]+1)),sd(log2(rsem_her2$ERBB2[13:17]+1)),5,mean(log2(rsem_her2$ERBB2[1:12]+1)),sd(log2(rsem_her2$ERBB2[1:12]+1)),12)
####Hedge's standardized mean/effect size using Rsubread pipeline 
standardized_mean(m.1=mean((log2(rsub_her2$ERBB2[13:17]+1))),sd.1=sd((log2(rsub_her2$ERBB2[13:17]+1))),n.1=5,m.2=mean((log2(rsub_her2$ERBB2[1:12]+1))),sd.2=sd((log2(rsub_her2$ERBB2[1:12]+1))),n.2 = 12)
```

Both methods find significant differences in the ERBB2 read counts between GFP control and HER2 samples. However, coefficient of variable is less with Rsubread data than TCGA pipeline processed data. Moreover, standardized mean differences in ERBB2 gene counts between HER2 activated and control samples using TCGA pipeline is smaller than the Rsubread pipeline processed data.

Using TCGA pipeline produced expected gene counts data, coefficient of variation in HER2 gene for GFP overexpressed control samples are: `r sd(rsem_her2_expected_counts$ERBB2[13:17])/mean(rsem_her2_expected_counts$ERBB2[13:17])` and HER2 overexpressed samples are: `r sd(rsem_her2_expected_counts$ERBB2[13:17])/mean(rsem_her2_expected_counts$ERBB2[13:17])`

Using Rsubreaad pipeline produced expected gene counts data, coefficient of variation in HER2 gene for GFP overexpressed control samples are: `r sd(rsub_her2$ERBB2[1:12])/mean(rsub_her2$ERBB2[1:12])` and HER2 overexpressed samples are: `r sd(rsub_her2$ERBB2[13:17])/mean(rsub_her2$ERBB2[13:17])`

### "Normalized gene counts" level analysis

We compared normalized gene counts results using GFP control protein and ERBB2 protein overexpressed human mammary epithelial cell samples. Log based 2 gene expression of the TCGA Level 3 data, Rsubread processed FPKM and TPM data are used for this analysis.

First, we computed the consistency of the gene expression across common genes after normalization. Second, empiric cumulative distribution for all samples from two pipelines. Second, we compared the total mapped read per HMEC samples from TCGA pipeline & Rsubread pipeline. Third, we compared HER2(ERBB2) gene expressions between control and HER2-overexpressed samples.

```{r echo=FALSE}
#######################comparing gene counts results ############
par( mfrow = c( 1,3 ) ,lwd=4)
TCGA_her2_filtered<-TCGA_her2[!duplicated(TCGA_her2$Gene),]
rownames(TCGA_her2_filtered)<-TCGA_her2_filtered$Gene
TCGA_her2<-subset(TCGA_her2_filtered,select=-Gene)
TCGA_her2_log2<-log2(subset(TCGA_her2_filtered,select=-Gene)+1)

###Coefficient of variation in GFP samples across all common genes
####Coefficient of variation in TCGA pipeline processed data
com_genes_TCGA<-TCGA_her2[rownames(TCGA_her2)%in%rownames(rsub_fpkm),]
hist(na.omit(apply(com_genes_TCGA,1,sd)/apply(com_genes_TCGA,1,mean)),main = "TCGA Level 3",xlab = "Coefficient of variation",ylim=c(0,12500),lwd=4,ylab="Number of genes", breaks = 20)
hist(na.omit(apply(com_genes_TCGA[,1:12],1,sd)/apply(com_genes_TCGA[,1:12],1,mean)),main = "TCGA Level 3",xlab = "Coefficient of variation",ylim=c(0,12500),lwd=4,ylab="Number of genes", breaks = 20)
print(paste("Coefficient of variation in TCGA Level 3 data across 19585 genes in the control samples:",median(na.omit(apply(com_genes_TCGA[,1:12],1,sd)/apply(com_genes_TCGA[,1:12],1,mean))),sep=" "))

hist(na.omit(apply(com_genes_TCGA[,13:17],1,sd)/apply(com_genes_TCGA[,13:17],1,mean)),main = "TCGA Level 3",xlab = "Coefficient of variation",ylim=c(0,12500),lwd=4,ylab="Number of genes", breaks = 20)
print(paste("Coefficient of variation in TCGA Level 3 data across 19585 genes in the HER2-overexpressed samples:",median(na.omit(apply(com_genes_TCGA[,13:17],1,sd)/apply(com_genes_TCGA[,13:17],1,mean))),sep=" "))

tcga_her2_normalized<-data.frame(t(TCGA_her2["ERBB2",]))


####Coefficient of variation in Rsubread pipeline processed data
com_genes_fpkm<-rsub_fpkm[rownames(rsub_fpkm)%in%rownames(com_genes_TCGA),]
hist(na.omit(apply(com_genes_fpkm[,13:17],1,sd)/apply(com_genes_fpkm[,13:17],1,mean)),main = "Rsubread FPKM",xlab = "Coefficient of variation",ylim=c(0,12500),lwd=4,ylab="Number of genes",breaks=20)

print(paste("Coefficient of variation in Rsubread FPKM normalized data across 19585 genes in the control samples:",median((na.omit(apply(com_genes_fpkm[,1:12],1,sd)/apply(com_genes_fpkm[,1:12],1,mean)))),sep=''))

print(paste("Coefficient of variation in Rsubread FPKM normalized data across 19585 genes in the HER2-overexpressed samples:",median((na.omit(apply(com_genes_fpkm[,13:17],1,sd)/apply(com_genes_fpkm[,13:17],1,mean)))),sep=''))


rsub_fpkmlog_her2<-data.frame(t(rsub_fpkmlog["ERBB2",]))
rsub_fpkm_her2<-data.frame(t(rsub_fpkm["ERBB2",]))
com_genes_tpm<-rsub_fpkm[rownames(rsub_tpm)%in%rownames(com_genes_TCGA),]

hist(na.omit(apply(com_genes_tpm[,13:17],1,sd)/apply(com_genes_tpm[,13:17],1,mean)),main = "Rsubread TPM",xlab = "Coefficient of variation",ylim=c(0,12500),lwd=4,ylab="Number of genes")

print(paste("Coefficient of variation in Rsubread TPM normalized data across 19585 genes in the control samples:",median((na.omit(apply(com_genes_tpm[,1:12],1,sd)/apply(com_genes_tpm[,1:12],1,mean)))),sep=''))

print(paste("Coefficient of variation in Rsubread TPM normalized data across 19585 genes in the HER2-overexpressed samples:",median((na.omit(apply(com_genes_tpm[,13:17],1,sd)/apply(com_genes_tpm[,13:17],1,mean)))),sep=''))

rsub_tpm_her2<-data.frame(t(rsub_tpm["ERBB2",]))
rsub_tpmlog_her2<-data.frame(t(rsub_tpmlog["ERBB2",]))

#######post normalization ecdf
ecdf_all_ex<-apply(log2(TCGA_her2+1),2,ecdf)
par( mfrow = c( 1, 3 ) )
plot(ecdf_all_ex[[1]],xlab=NA, ylab = NA,col="blue",main="TCGA Level 3",ylim=c(0,1),xlim = c(0,20),cex.axis=1.5, cex.lab=1.5,)
for(i in 2:12){lines(ecdf_all_ex[[i]],xlab=NA, ylab = NA,col="blue")}
for(i in 13:17){lines(ecdf_all_ex[[i]],xlab=NA, ylab = NA,col="brown")}
###using Rsubread pipeline aligned data
ecdf_all<-apply(rsub_fpkmlog,2,ecdf)
plot(ecdf_all[[1]],col="blue",main="Rsubread FPKM",ylim=c(0,1),xlim = c(0,20),cex.axis=1.5, cex.lab=1.5,xlab="log2(normalized expression)",ylab="Cumulative proportion")
for(i in 2:12){lines(ecdf_all[[i]],xlab=NA,ylab = NA,col="blue")}
for(i in 13:17){lines(ecdf_all[[i]],xlab=NA,ylab = NA,col="brown")}

ecdf_all_t<-apply(rsub_tpmlog,2,ecdf)
plot(ecdf_all_t[[1]],col="blue",main="Rsubread TPM",ylim=c(0,1),xlim = c(0,20),cex.axis=1.5, cex.lab=1.5,xlab="log2(normalized expression)",ylab="Cumulative proportion")
for(i in 2:12){lines(ecdf_all_t[[i]],xlab=NA,ylab = NA,col="blue")}
for(i in 13:17){lines(ecdf_all_t[[i]],xlab=NA,ylab = NA,col="brown")}


###Creating boxplots of the normalized ERBB2 expression   
par( mfrow = c( 1, 1 ) )
par(mar=c(5, 4.5, 3.5, 0.5)) 
boxplot(log2(tcga_her2_normalized$ERBB2[1:12]+1),log2(tcga_her2_normalized$ERBB2[13:17]+1),rsub_fpkmlog_her2$ERBB2[1:12],rsub_fpkmlog_her2$ERBB2[13:17],rsub_tpmlog_her2$ERBB2[1:12],rsub_tpmlog_her2$ERBB2[13:17],ylab="log2(HER2 gene expression values)",main="Comparing HER2 normalized expression between\n control and her2 samples",range=0,cex.axis=1.5, cex.lab=1.5,outpch=NA,col='grey75',medcol="grey75",lwd=4,border = "grey35")
names=c("TCGA\nGFP","TCGA\nHER2","Rsubred FPKM\nGFP", "Rsubred FPKM\nHER2","Rsubred TPM\nGFP", "Rsubred TPM\nHER2")
text(seq(1,6,by=1),par("usr")[3] - 2, labels = names, srt = 45, pos = 1, xpd = TRUE)
points(jitter(rep(1,12),factor=2),log2(tcga_her2_normalized$ERBB2[1:12]+1),pch=4,cex=2,cex.lab=1.5)
points(jitter(rep(2,5),factor=2),log2(tcga_her2_normalized$ERBB2[13:17]+1),pch=4,cex=2,cex.lab=1.5)
points(jitter(rep(3,12),factor=2),rsub_fpkmlog_her2$ERBB2[1:12],pch=4,cex=2,cex.lab=1.5)
points(jitter(rep(4,5),factor=2),rsub_fpkmlog_her2$ERBB2[13:17],pch=4,cex=2,cex.lab=1.5)
points(jitter(rep(5,12),factor=2),rsub_tpmlog_her2$ERBB2[1:12],pch=4,cex=2,cex.lab=1.5)
points(jitter(rep(6,5),factor=2),rsub_tpmlog_her2$ERBB2[13:17],pch=4,cex=2,cex.lab=1.5)


###t.test to see if there is significance
t.test(log2(tcga_her2_normalized$ERBB2[1:12]+1),log2(tcga_her2_normalized$ERBB2[13:17]+1))
t.test(rsub_fpkmlog_her2$ERBB2[1:12],rsub_fpkmlog_her2$ERBB2[13:17])
t.test(rsub_tpmlog_her2$ERBB2[1:12],rsub_tpmlog_her2$ERBB2[13:17])

###Standardized mean difference: TCGA pipeline normalized ERBB2 expression values
standardized_mean(m.1=mean((log2(tcga_her2_normalized$ERBB2[13:17]+1))),sd.1=sd((log2(tcga_her2_normalized$ERBB2[13:17]+1))),n.1=5,m.2=mean((log2(tcga_her2_normalized$ERBB2[1:12]+1))),sd.2=sd((log2(tcga_her2_normalized$ERBB2[1:12]+1))),n.2=12)

###Standardized mean difference: Rsubread pipeline FPKM normalized ERBB2 expression values
standardized_mean(mean(rsub_fpkmlog_her2$ERBB2[13:17]),sd(rsub_fpkmlog_her2$ERBB2[13:17]),5,mean(rsub_fpkmlog_her2$ERBB2[1:12]),sd(rsub_fpkmlog_her2$ERBB2[1:12]),12)

###Standardized mean difference:Rsubread pipeline TPM normalized ERBB2 expression values
standardized_mean(mean(rsub_tpmlog_her2$ERBB2[13:17]),sd(rsub_tpmlog_her2$ERBB2[13:17]),5,mean(rsub_tpmlog_her2$ERBB2[1:12]),sd(rsub_fpkmlog_her2$ERBB2[1:12]),12)


```
Coefficient of variation in ERBB2 gene in HER2 overexpressed samples using TCGA pipeline is: `r sd(tcga_her2_normalized$ERBB2[13:17])/mean(tcga_her2_normalized$ERBB2[13:17])`
using Rsubread FPKM pipeline is: `r sd(rsub_fpkm_her2$ERBB2[13:17])/mean(rsub_fpkm_her2$ERBB2[13:17])`and using Rsubread TPM pipeline is: `r sd(rsub_tpm_her2$ERBB2[13:17])/mean(rsub_tpm_her2$ERBB2[13:17])`

###Assessing the number of zeroes per sample
Now checking the total number of zeros per samples in the common  samples accross PanCan12 and our datasets
```{r echo=FALSE}
colnames(pancan12_zero)<-"PANCAN12"
colnames(pancan20_tpm_zero)<-"TPM"
all_zeros<-merge_drop(pancan12_zero,pancan20_tpm_zero)
#3380 samples are common
par(mfrow = c(1, 2),lwd=4)
h1<-hist(all_zeros$PANCAN12,xlab='',ylab='',main='',xlim=c(0,8000),ylim=c(0,800),lwd=4,breaks = 25)
abline(v=median(all_zeros$PANCAN12),col="red",lty=2)
h2<-hist(all_zeros$TPM,xlab='',ylab='',main='',xlim=c(0,8000),ylim=c(0,800),lwd=4,breaks=25)
abline(v=median(all_zeros$TPM),col="red",lty=2)
t.test(all_zeros$PANCAN12,all_zeros$TPM)
```

### Comparing Her2 predictions

We predicted HER2 pathway using our overexpressed samples in human mammary epithelial cells(HMECs) in TCGA breast cancer samples. We used TCGA pipeline processed HMEC datasets on TCGA pipeline processed TCGA BRCA samples and Rsubread pipelined processed HMEC datasets on Rsubread processed TCGA BRCA samples. We only compared BRCA samples that are common and we have her2 status confirmed by immunohistochemistry. Our goal here is to compare how well these two pipeline and normalization method differenciate HER2 pathway activity predictions where we know the status of HER2 overexpression.

We removed all the genes that had zero expression accross all HMEC samples and TCGA samples. Then, we used quantile normalization method to adjust for batch variation between our HMEC training samples and TCGA BRCA samples. We used binary regression model developed by West et al to generate HER2 signatures using top 200 genes. The signature generated from each pipeline processed data was used to predict on TCGA BRCA RNA-Seq sample processed by the same pipeline. Both FPKM-log2 and TPM-log2 normalized Rsubread processed predictions were compiled to one file and TCGA upper-quantile log-2 normalized pipeline procssed predictions were extracted to another file for analysis.

```{r echo=FALSE}
#############Predicted HER2 pathway activity analysis#############################################

all_preds<-merge_drop(rsub_preds,tcga_preds,by=0)
brca_clinical<-subset(clinicals,clinicals[,'tumor_tissue_site']=='Breast',select=c("bcr_patient_barcode","her2_status_by_ihc"))
common_all<-merge_drop(all_preds,brca_clinical,by=0)
all_preds_pos_neg<-subset(common_all,common_all$her2_status_by_ihc=="Negative"|common_all$her2_status_by_ihc=="Positive")
all_ranked<-apply(all_preds_pos_neg[,1:3],2,rank)
all<-cbind(all_ranked,all_preds_pos_neg[,4:5])
ihc_neg<-subset(all,all$her2_status_by_ihc=="Negative")
ihc_pos<-subset(all,all$her2_status_by_ihc=="Positive")

```
`r length(rownames(ihc_pos))` HER2 negative and `r length(rownames(ihc_pos))` HER2 positive samples are being compared here. 

We compared the boxplots for HER2 pathway activation status predictions between HER2(+) and HER2(-) BRCA samples.

```{r echo=FALSE}
##############boxplot of ranked estimated HER2 pathway activity
##in TCGA BRCA samples####
par(mfrow = c(1, 1))
par(mar=c(5, 4.6, 2.5, 0.6)) # figure margins
boxplot(ihc_pos$Rsem_log_q_200_f,ihc_neg$Rsem_log_q_200_f,ihc_pos$FPKM_log_q_200_f,ihc_neg$FPKM_log_q_200_f,ihc_pos$TPM_log_q_200_f,ihc_neg$TPM_log_q_200_f,cex.axis=1.5, cex.lab=1.5,outpch=NA,range=0,cex.axis=1, cex.lab=0.7,outpch=NA,col='grey75',medcol="grey5",lwd=4,border = "grey5", main="Comparison of rank-based estimate \nof HER2 activation",ylab="Ranked HER2 prediction")
names=c("TCGA\nLevel3\nHER2(+)","TCGA\nLevel3\nHER2(-)","Rsubred\nFPKM\nHER2(+)", "Rsubred\nFPKM\nHER2(-)","Rsubred\nTPM\nHER2(+)", "Rsubred\nTPM\nHER2(-)")
text(seq(1,6,by=1),par("usr")[3] - 4.5, labels = names, srt = 45, pos = 1, xpd = TRUE)
ihc_neg_t<-subset(common_all,common_all$her2_status_by_ihc=="Negative")
ihc_pos_t<-subset(common_all,common_all$her2_status_by_ihc=="Positive")

##coefficient of variation in TCGA pipeline processed HER2 predictions
print(paste("Coefficient of variation in TCGA pipeline processed HER2 predictions in HER2(-) BRCA samples",sd(ihc_neg_t$Rsem_log_q_200_f)/mean(ihc_neg_t$Rsem_log_q_200_f),sep=' '))
print(paste("Coefficient of variation in TCGA pipeline processed HER2 predictions in HER2(+) BRCA samples",sd(ihc_pos_t$Rsem_log_q_200_f)/mean(ihc_pos_t$Rsem_log_q_200_f),sep=' '))

##coefficient of variation in Rsubread FPKM pipeline processed HER2 predictions
print(paste("Coefficient of variation in Rsubread FPKM processed HER2 predictions in HER2(-) BRCA samples",sd(ihc_neg_t$FPKM_log_q_200_f)/mean(ihc_neg_t$FPKM_log_q_200_f),sep="  "))
print(paste("Coefficient of variation in Rsubread FPKM processed HER2 predictions in HER2(+) BRCA samples",sd(ihc_pos_t$FPKM_log_q_200_f)/mean(ihc_pos_t$FPKM_log_q_200_f),sep="  "))
##coefficient of variation in Rsubread TPM pipeline processed HER2 predictions
print(paste("Coefficient of variation in Rsubread TPM  processed HER2 predictions in HER2(-) BRCA samples",sd(ihc_neg_t$TPM_log_q_200_f)/mean(ihc_neg_t$TPM_log_q_200_f),sep="  "))
print(paste("Coefficient of variation in Rsubread TPM processed HER2 predictions in HER2(+) BRCA samples",sd(ihc_pos_t$TPM_log_q_200_f)/mean(ihc_pos_t$TPM_log_q_200_f),sep="  "))

##Calculating standardized mean differences between the HER2(+) and HER2(-) groups
print(paste("Standardized mean difference in predicrion between HER2 (+) and HER2 (-) samples for TCGA Level 3 data :",standardized_mean(m.1=mean(ihc_pos_t$Rsem_log_q_200_f),sd.1=sd(ihc_pos_t$Rsem_log_q_200_f),n.1=length(ihc_pos_t$Rsem_log_q_200_f),m.2=mean(ihc_neg_t$Rsem_log_q_200_f),sd.2=sd(ihc_neg_t$Rsem_log_q_200_f),n.2=length(ihc_neg_t$Rsem_log_q_200_f)),sep=' '))

print(paste("Standardized mean difference in predicrion between HER2 (+) and HER2 (-) samples for Rsubread FPKM data :",standardized_mean(m.1=mean(ihc_pos_t$FPKM_log_q_200_f),sd.1=sd(ihc_pos_t$FPKM_log_q_200_f),n.1=length(ihc_pos_t$FPKM_log_q_200_f),m.2=mean(ihc_neg_t$FPKM_log_q_200_f),sd.2=sd(ihc_neg_t$FPKM_log_q_200_f),n.2=length(ihc_neg_t$FPKM_log_q_200_f)),sep=' '))

print(paste("Standardized mean difference in predicrion between HER2 (+) and HER2 (-) samples for FPKM TPM data :",standardized_mean(m.1=mean(ihc_pos_t$TPM_log_q_200_f),sd.1=sd(ihc_pos_t$TPM_log_q_200_f),n.1=length(ihc_pos_t$TPM_log_q_200_f),m.2=mean(ihc_neg_t$TPM_log_q_200_f),sd.2=sd(ihc_neg_t$TPM_log_q_200_f),n.2=length(ihc_neg_t$TPM_log_q_200_f)),sep=' '))

## t-tests comparing HER(+) and HER(-) prediction
t.test(ihc_pos_t$Rsem_log_q_200_f,ihc_neg_t$Rsem_log_q_200_f)# For TCGA Level 3: p-value = 2.009e-05
t.test(ihc_pos_t$FPKM_log_q_200_f,ihc_neg_t$FPKM_log_q_200_f)#For Rsubread FPKM: p-value = 1.493e-10
t.test(ihc_pos_t$TPM_log_q_200_f,ihc_neg_t$TPM_log_q_200_f)#For Rsubread TPM:p-value = 3.197e-12

```

Both pipelines were significant in differentiating between HER2(+) and HER(-) samples using HER2 signature. However,  Rsubread has the bigger standardized mean difference between the two subtypes.

The consistency of the predictions was assessed. TCGA's pipeline produced predictions with higher coeffecient of variation: 
HER2(+) samples `r sd(ihc_pos_t$Rsem_log_q_200_f)/mean(ihc_pos_t$Rsem_log_q_200_f)` and HER2(-) samples`r sd(ihc_neg_t$Rsem_log_q_200_f)/mean(ihc_neg_t$Rsem_log_q_200_f)` compared to Rsubread's coefficient of variation.  
For FPKM data: HER2(+) `r sd(ihc_pos_t$FPKM_log_q_200_f)/mean(ihc_pos_t$FPKM_log_q_200_f)` and HER2(-)`r sd(ihc_neg_t$FPKM_log_q_200_f)/mean(ihc_neg_t$FPKM_log_q_200_f)`  
For TPM data: HER2(+) `r sd(ihc_pos_t$TPM_log_q_200_f)/mean(ihc_pos_t$TPM_log_q_200_f)` and HER2(-) `r sd(ihc_neg_t$TPM_log_q_200_f)/mean(ihc_neg_t$TPM_log_q_200_f)`.

Finally, we are assessing the classification predictions using 10-fold cross validation  and using TCGA Level 3 and Rsubread TPM normalized lung adenocarcinoma and lung squamous carcinoma samples. Rsubread processed data had better accuracy in predicting class than TCGA Level 3 data.
```{r echo=FALSE}
par(mfrow = c(1, 1),lwd=4)
actual12 = data12$ActualClass
predictions12 = data12$LUAD_Probability
auc = plotROC(actual12, predictions12, TRUE)
title("TCGA Level 3 LUAD vs LUSC")
actual20 = data20$ActualClass
predictions20 = data20$LUAD_Probability
auc = plotROC(actual20, predictions20, TRUE)
title("Rsubread TPM LUAD vs LUSC")
```


```{r echo=FALSE}
time<-format(Sys.time(),"%a %b %d %X %Y")
```
This analysis was run on `r time` 
